### 2013-05-08 writeup/paper.tex

\documentclass[10pt, conference, compsocconf]{IEEEtran}

%% Custom TeX

\usepackage{listings}
\usepackage{xcolor}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{rotating}
\usepackage{amssymb}
\usepackage{url}
\usepackage[splitrule]{footmisc}

%% for comments
\newcommand{\comment}[3][\color{red}]{{#1{[{#2}: {#3}]}}}
\newcommand{\code}[1]{\textsf{\small #1}}
\newcommand{\kris}[1]{\comment[\color{orange}]{km}{#1}}

\newcommand{\thickhline}{\noalign{\hrule height 1pt}}

%% End of custom TeX

\begin{document}
%
% --- Author Metadata here ---
% --- End of Author Metadata ---

\title{Using Big Data for Music Synthesis}

%%\numberofauthors{3}
\author{
\IEEEauthorblockN{Christopher Imbriano, Amanda Strickler, Kristopher Micinski
\IEEEauthorblockA{Computer Science Department, University of Maryland,
  College Park, MD 20742, USA\\
  \{micinski, pdphelp, jfoster\}@cs.umd.edu}
}

\maketitle

\begin{abstract}
\end{abstract}

\section{Introduction}
\label{sec:introduction}

The big data revolution has revolutionized the way we think about,
design, and implement algorithms at scale.

We present a system to synthesize music using large amounts of data.
The input is a large corpus of images, which is subsequently processed
by our system, and the result is a musical score.

We study the music we

\section{Overview}
\label{sec:overview}

The core idea in converting image data to music is to view the
information in an image (color, size, opacity, etc..) as a mapping to
notes and durations with which those notes are sustained.  In our
system we chose MIDI as the output format.  Our system can then be
modeled as a function \kris{transducer?} from image data to MIDI
output.  Note that the choice of MIDI limits the expressivity of our
music.  Going to other formats (e.g., MP3, OOG, etc..) could provide a
larger range of primitives with which to construct music, but does not
enjoy the simplicity that MIDI provides.

Our system is modeled as a function from ordered sets of images to a
MIDI file.  In our view, a MIDI file consists of note, duration pairs.

%% \[
%%     f : image^{\ast} \to (
%% \]

\paragraph{The HSB format}

To convert image data to musical notes, we need to establish a
bijection between the data in the image and the range of music in the
song.  The way we do this is inspired by viewing colors in images as
HSV values.  In HSV, each color has three components: a hue, a
saturation, and a value.  These components are visualized (as shown in
Figure~\ref{fig:hsv}) using a wheel.  The components represent:

\begin{describe}
\item[Hue] The actual ``color'' that would be described.  At full
  saturation and value, the hue determines the color seen.  This is
  visualized as going around a wheel of colors.

\item[Saturation] The degree to which the color is saturated from
  white.  Zero saturation is a blank (white) surface, while increasing
  values show more permeance of color.

\item[Value] The amount of light in the color.  Colors with low values
  are dark, and those with higher values are brighter.  This is
  visualized by going up the cylinder.
\end{describe}

\kris{Here I'm going to draw a diagram which shows two points on the
  HSV wheel and a musical clef, illustrating the position.}

\begin{figure}
  \centering
  \includegraphics{hsv}
  \caption{The HSV color cylinder}
  \label{fig:hsv}
\end{figure}

\paragraph{Generating music}

To generate music, we take each image in our corpus and view it as a
single step in time.  Musical scores are made by streaming large
amounts of images.  (In fact, one possible idea is to use image data
from subsequent frames in movies to get streams of related images.)
Because our music would be relatively uninteresting if it had only one
``player'', we also include several \emph{regions} in the generated
music.  Each region is a distinct musical voice that can play
different notes and is mixed together with the rest of the music.  To
state this formally, each image is mapped to a set of regions and
notes:

\[
    f : A_{i,j} \to note^n
\]

where $i$ and $j$ are the dimensions of the image, and $n$ is the
number of distinct regions of the image.  The obvious idea is to
divide the image into $n$ distinct quadrants of equal size and project
the colors in that region to notes.  One problem is that we often deal
with relatively large images: if we have a smaller set of quadrants
how do we take this to a HSV value?  We solve this problem by taking
thumbnails of images, essentially doing a projection from $A_{i,j}$
down to $A_{n^{.5},n^{.5}}$.  This also tacitly stipulates that we
start out with square images (otherwise we lose some of the image data
in the projection): we handle this step as a first preprocessing
phase.

\section{Implementation}

\begin{figure}
  \centering
  \caption{High level view of the pipeline used for the project}
  \label{fig:project-pipeline}
\end{figure}

Our implementation consists of three MapReduce phases, tied together
with helper scripts.  The high level pipeline is shown in
Figure~\ref{fig:project-pipeline}.

\paragraph{Conversion to intermediate representation}
\label{sec:conversion}

Image data comes in a variety of different formats: JPEG, GIF, PNG,
etc..  Dealing with the plethora of formats presents a problem for our
translation because there are many artifacts of the formats
(annotations, geotagging, compression, etc..) which are unnecessary
for our implementation.  To keep the core abstraction of images as
arrays of bytes (the abstraction described in
Section~\ref{sec:overview}) the first stage of our Hadoop
implementation\footnote{\code{src/main/phase1}} converts the surface
image representation to the sequence file format used in subsequent
processing.  Although it may appear inefficient to hold an
uncompressed image in memory, this is actually ideal for Hadoop:
because computing cycles (incurred through recompressing and sending
the images) are the cost, compared with network traffic \kris{does
  this make sense???}.

The conversion in our prototype uses the java \code{ImageBuffer} class
to do image input and conversion.  This class holds requires a large
amount of memory to represent image objects, however this is fine
because we hold only a small number of images in memory at any one
time (using statics) when mapping over our input corpus.

\paragraph{Translating to note, region pairs}

After taking the image data to an intermediate format, we must map the
image data to a set of note / region pairs for use in the MIDI.  To do
this, we do projection

\section{Results}

\subsection{Sample datasets}

To test the system, we use a variety of different sets of images.
These images were obtained from
Flickr\footnote{\url{http://www.flickr.com/}}, using a public API and
a Python script \cite{flickerpy} \kris{Did we actually use this?  Did
  Amanda write the code?}

\section{Related work}

Many domain specific programming languages exist for music synthesis
[\cite{haskore},\cite{lilypad}].  Many of these languages synthesize
music in the small ().  By contrast, our approach focuses on using
large amounts of data to create music...

\section{Conclusion}

\bibliographystyle{IEEEtran}
\bibliography{paper}

\end{document}

### 2013-05-08 writeup/paper.tex
pushing more of the report

@@
 \usepackage{listings}
-\usepackage{xcolor}
 \usepackage{xspace}
@@
 \usepackage[splitrule]{footmisc}
+
+\usepackage[dvipsnames]{xcolor}
+
+\usepackage{listings}
+\lstloadlanguages{Ruby}
+\lstset{
+basicstyle=\ttfamily\color{black},
+commentstyle = \ttfamily\color{red},
+keywordstyle=\ttfamily\color{blue},
+stringstyle=\color{orange}}
 
@@
 \author{
-\IEEEauthorblockN{Christopher Imbriano, Amanda Strickler, Kristopher Micinski
+\IEEEauthorblockN{Christopher Imbriano, Amanda Strickler, Kristopher Micinski}
 \IEEEauthorblockA{Computer Science Department, University of Maryland,
-  College Park, MD 20742, USA\\
-  \{micinski, pdphelp, jfoster\}@cs.umd.edu}
+  College Park, MD 20742, USA
+}
 }
@@
 The big data revolution has revolutionized the way we think about,
-design, and implement algorithms at scale.
+design, and implement algorithms at scale.  As such, problems from all
+different areas of computer science have been lifted into the
+MapReduce framework.  By contrast, MapReduce has been used relatively
+less to explore art or creative tasks.  Our project looks to answer:
+what kinds of art can we create harnessing large amounts of data at
+scale?
 
@@
 The input is a large corpus of images, which is subsequently processed
-by our system, and the result is a musical score.
+by our system, and the result is a musical score.  We then mix and
+listen to the music.
 
-We study the music we
+\paragraph{Motivation}
+
+We are currently unaware of any projects using MapReduce to create art
+or any other representation of digital media.  Because of this, we
+anticipate that there might be problems in handling media (buffering,
+representing, etc..).  We anticipate that exploring this area might
+open up potential applications.
 
@@
 Our system is modeled as a function from ordered sets of images to a
-MIDI file.  In our view, a MIDI file consists of note, duration pairs.
+MIDI file.  In our view, a MIDI file consists of note and duration
+pairs.  Taken together, these constitute a score by chaining together
+multiple pairs over the course of time:
+
+\[
+MIDI : (Note,Duration)^{\ast}
+\]
 
@@
 
-\paragraph{The HSB format}
+\paragraph{The HSV format}
 
@@
 
-\begin{describe}
-\item[Hue] The actual ``color'' that would be described.  At full
+\begin{enumerate}
+\item Hue --- The actual ``color'' that would be described.  At full
   saturation and value, the hue determines the color seen.  This is
@@
 
-\item[Saturation] The degree to which the color is saturated from
+\item Saturation --- The degree to which the color is saturated from
   white.  Zero saturation is a blank (white) surface, while increasing
@@
 
-\item[Value] The amount of light in the color.  Colors with low values
-  are dark, and those with higher values are brighter.  This is
+\item Value --- The amount of light in the color.  Colors with low
+  values are dark, and those with higher values are brighter.  This is
   visualized by going up the cylinder.
-\end{describe}
+\end{enumerate}
 
@@
 
+In our transformation to music, we map colors in HSV to notes in MIDI.
+To do this, we map colors in RGB to HSV using the \code{\#RGBtoHSV}
+method of the \code{java.awt.Color} class.
+
 \begin{figure}
   \centering
-  \includegraphics{hsv}
+  \includegraphics[width=.4\textwidth]{hsv.png}
   \caption{The HSV color cylinder}
@@
 \paragraph{Generating music}
+\label{sec:generating}
 
@@
 start out with square images (otherwise we lose some of the image data
-in the projection): we handle this step as a first preprocessing
-phase.
+in the projection): we handle this step as a first preprocessing phase
+where we take images and project them to square images.
+
+In the case of MIDI, notes are pairs of integers in the range
+$[-1,127]$.  The first element of the pair represents the \emph{tone}
+, and the second element of the pair represents the \emph{velocity}
+(how loud the note is):
+
+\[
+    note : tone \times velocity
+\]
+
+where a velocity of $-1$ represents the lack of any tone, and $127$
+represents the loudest possible tone.  Colors start at $C_0$ and move
+up over ten octaves.
+
+\paragraph{Conversion to colors}
 
@@
 
-\begin{figure}
+\begin{figure*}
   \centering
+  \includegraphics[width=.85\textwidth]{pipeline.png}
   \caption{High level view of the pipeline used for the project}
   \label{fig:project-pipeline}
+\end{figure*}
+
+Our implementation consists of scripts to transform the input,
+MapReduce phases, and a music mixing tool.  The high level pipeline is
+shown in Figure~\ref{fig:project-pipeline}.
+
+\paragraph{Conversion of images to thumbnails}
+
+Conversion to music happens by doing operations on colors and image
+data.  To do the conversion to tone, velocity pairs, we transform
+colors in the input data, meaning that we have to start with a
+representation of $R$ colors for each image.  To do this
+transformation, we considered using a MapReduce job: mapping over the
+sets of images in our input corpus.  Ultimately we found that
+converting images to thumbnails was better accomplished using scripts:
+which allowed us to do other manipulation as well.  We anticipate that
+if this became a problem we could have moved the image processing code
+to Java and pass in our input corpus to the initial stage of
+MapReduce.  Our current implementation uses the RMagick \cite{rmagick}
+gem, which is a Ruby library interfacing to ImageMagick
+\cite{imagemagick}.  ImageMagick includes facilities for taking images
+to thumbnails.
+
+\begin{figure}
+\begin{lstlisting}[language=Ruby]
+require 'rmagick'
+
+class ThumbMaker
+  # Convert set of images to thumbnails
+  def self.convert_thumbs(dir_path, size)
+      # ... 
+      puts "Making thumbnail for..."
+      Dir.glob("#{path}/*.jpg") do |fname|
+        # ...
+        fname = # ...
+        img = Magick::Image.read(fname)[0]
+        thumb = img.thumbnail(size, size)
+        thumb.write(thumb_dest)
+      end
+    end
+end
+
+\end{lstlisting}
+\caption{Relevant code from the module used for image to thumbnail
+  conversion}
 \end{figure}
-
-Our implementation consists of three MapReduce phases, tied together
-with helper scripts.  The high level pipeline is shown in
-Figure~\ref{fig:project-pipeline}.
 
@@
 because computing cycles (incurred through recompressing and sending
-the images) are the cost, compared with network traffic \kris{does
-  this make sense???}.
+the images) are the cost, compared with network traffic.
 
@@
 
+\paragraph{Mixing and producing playable music}
+
+The result of our mapreduce pipeline is a set of MIDI music: one for
+each region in the image set.  To turn these into music that can be
+played, we use mixing software (such as GarageBand) to
+mix the regions into full scores.
+
 \section{Results}
@@
 Flickr\footnote{\url{http://www.flickr.com/}}, using a public API and
-a Python script \cite{flickerpy} \kris{Did we actually use this?  Did
-  Amanda write the code?}
+a Python script \cite{flickerpy} to get test data.
 

### 2013-05-09 writeup/paper.tex
adding some more information relating to secondary sorting

@@
 \usepackage{listings}
-\lstloadlanguages{Ruby}
+\lstloadlanguages{Ruby,Java}
 \lstset{
@@
 
-\paragraph{Motivation}
+\subsection{Motivation}
 
@@
 
-\paragraph{The HSV format}
+\subsection{The HSV format}
 
@@
 
-\paragraph{Generating music}
+\subsection{Generating music}
 \label{sec:generating}
@@
 
-\paragraph{Conversion to colors}
+We represent each region of the image by a representative color, which
+is obtained (as noted previously) using thumbnailing.
+
+\begin{figure}
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.4\linewidth]{image1}
+  \caption{A subfigure}
+  \label{fig:sub1}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.4\linewidth]{image1}
+  \caption{A subfigure}
+  \label{fig:sub2}
+\end{subfigure}
+\caption{A figure with two subfigures}
+\label{fig:test}
+\end{figure}
 
@@
 
+Our implementation for this conversion happens in the
+\code{WriteImagesToSequenceFile} class.  They key novelty in this
+class is the conversion of image data to a \code{SequenceFile} based
+representation.  A \code{SequenceFile} works well for our application
+because we need to stream potentially large amounts of image data (raw
+bytes).  This is fundamentally no different than creating a custom
+Writable instance to serialize to a \code{SequenceFile} at a later
+point, but given the simplicity of our representation we say no need
+to do this.
+
+\begin{figure}
+\begin{lstlisting}[language=Java]
+public class WriteImagesToSequenceFile {
+  BufferedImage img = null;
+  SequenceFile.Writer writer = null;
+  FileSystem fs = FileSystem.get(conf);
+
+  FileStatus[] fss = 
+    fs.globStatus(new Path(inputPath 
+      + "/*.jpg"));
+
+  writer = SequenceFile.createWriter(fs, 
+    conf, outPath, IntWritable.class,
+    ArrayListOfIntsWritable.class);
+  int counter = 0;
+  for (FileStatus status : fss) {
+    IntWritable image_counter = 
+      new IntWritable();
+    ArrayListOfIntsWritable 
+      pixelData = 
+        new ArrayListOfIntsWritable();
+
+    img = ImageIO.read(
+      new File(status.getPath()
+               .toString().substring(5)));
+
+    // Get the RBG integer for each pixel 
+    // in 3X3
+    for(int j = 0; j < thumbs; j++) {
+      for(int i = 0; i< thumbs; i++) {
+        pixelData.add(img.getRGB(i,j));
+      }
+    }
+
+    // Write to the sequence file
+    image_counter.set(counter);
+    writer.append(image_counter, pixelData);
+    counter += 1;
+  } 
+}
+\end{lstlisting}
+\caption{Snippet of code from \code{WriteToSequenceFile} showing
+  format used in sequence files throughout the MapReduce jobs.}
+\end{figure}
+
 \paragraph{Translating to note, region pairs}
@@
 image data to a set of note / region pairs for use in the MIDI.  To do
-this, we do projection
+this, we map a conversion function across each of the different
+portions of each image.  The conversion happens by mapping over our
+set of images in the \code{Image2MusicMR} class.  The output of the
+image to music conversion is $n$ sets of note lists (where notes are
+expressed as pairs).
 
-\paragraph{Mixing and producing playable music}
+The strategy used in the conversion is to map over each pixel in an
+image, apply the conversion function, and emit it to a reducer to
+collate together.  After mapping over each pixel, we need to end up
+with a stream of images for each region: a naive implementation simply
+maps each pixel in our input corpus to these tone, velocity pairs:
+
+\[
+  map : id \times (r,g,b) \to region \times (tone \times velocity)
+\]
+
+However, when we actually mix the music, we need each of the streams
+to be in an ordered format.  Because the default sorting order does
+not take this into account, the music ends up jumbled (each of the
+regions is permuted with respect to input ID in an arbitrary way).  To
+fix this, we perform secondary sorting on the input ID:
+
+\[
+  map : id \times (r,g,b) \to (id \times region) \times (tone \times velocity)
+\]
+
+\subsection{Mixing and producing playable music}
 
@@
 each region in the image set.  To turn these into music that can be
-played, we use mixing software (such as GarageBand) to
-mix the regions into full scores.
+played, we use mixing software (such as GarageBand) to mix the regions
+into full scores.
 

### 2013-05-09 src/main/phase2/Color2Music.java
Adding space optimization: instead of creating new  object, convert to HSV inline: this also avoids going through java.awt as much

@@ -35,13 +35,44 @@ public class Color2Music {
   public static final int[] regionToMidiNote = {C, G, D, A, E, B, 
     F_SHARP, C_SHARP, G_SHARP, D_SHARP, A_SHARP, F};
 
-  public static MidiNote convert(Color c){
+  private static float[] hsvConversionArray = new float[3];
+  
+  /**
+   * Convert an RGB pixel to an HSV int array
+   * @param pixel
+   */
+  private static void RGBtoHSV(int pixel) {
+    float r = (pixel & 0xFF)/255.0F;
+    float g = (pixel >> 8)/255.0F;
+    float b = ((pixel >> 16) & 0xFF)/255.0F;
     
-    float[] hsvals = Color.RGBtoHSB(c.getRed(), c.getGreen(), c.getBlue(), null);
+    float min = Math.min(r,Math.min(g,b));
+    float max = Math.max(r,Math.max(g,b));
     
-    float hue = hsvals[0];
-    float saturation = hsvals[1];
-    float brightness = hsvals[2];
+    if (Math.min(r,Math.min(g,b))==Math.max(r,Math.max(g,b))) {
+     hsvConversionArray[0] = 0;
+     hsvConversionArray[1] = 0;
+     hsvConversionArray[2] = Math.min(r,Math.min(g,b));
+     return;
+    }
+
+    // Colors other than black-gray-white:
+    float d = (r==min) ? g-b : ((b==min) ? r-g : b-r);
+    float h = (r==min) ? 3 : ((b==min) ? 1 : 5);
+    hsvConversionArray[0] = 60*(h - d/(max - min));
+    hsvConversionArray[1] = (max - min)/max;
+    hsvConversionArray[2] = max;
+    return;
+  }
+  
+  public static MidiNote convert(int pixel){
+
+    
+    RGBtoHSV(pixel);
+    
+    float hue = hsvConversionArray[0];
+    float saturation = hsvConversionArray[1];
+    float brightness = hsvConversionArray[2];
     
     int tone = getMidiToneFromHSValues(hue, saturation, brightness);
     

### 2013-05-09 src/main/phase2/Image2MusicMR.java

@@ -63,7 +63,7 @@ public class Image2MusicMR extends Configured implements Tool {
         if(regionCounter > 9) LOG.info("ERROR: Region Counter > 9:" + regionCounter);
 
         pixel = pixels.get(i);
-        midiNote = Color2Music.convert(new Color(pixel));
+        midiNote = Color2Music.convert(pixel);
 
         //IMAGE_REGION.set(regionCounter);
         REGION_IMAGENO.set(regionCounter, key.get());

### 2013-05-10 writeup/paper.tex
adding paper

@@
 \usepackage{rotating}
+\usepackage{geometry}
 \usepackage{amssymb}
@@
 
-\title{Using Big Data for Music Synthesis}
+\title{MapReduce to My Ears\\
+{\Large Using Big Data for Music Synthesis}
+}
 
@@
 system we chose MIDI as the output format.  Our system can then be
-modeled as a function \kris{transducer?} from image data to MIDI
-output.  Note that the choice of MIDI limits the expressivity of our
-music.  Going to other formats (e.g., MP3, OOG, etc..) could provide a
-larger range of primitives with which to construct music, but does not
-enjoy the simplicity that MIDI provides.
+modeled as a function from image data to MIDI output.  Note that the
+choice of MIDI limits the expressivity of our music.  Going to other
+formats (such as MP3 or OOG) could provide a larger range of
+primitives with which to construct music, but does not enjoy the
+simplicity that MIDI provides.
 
-Our system is modeled as a function from ordered sets of images to a
-MIDI file.  In our view, a MIDI file consists of note and duration
-pairs.  Taken together, these constitute a score by chaining together
-multiple pairs over the course of time:
+Our system takes in ordered sets of images and produces a MIDI file.
+In our view, a MIDI file consists of note and duration pairs.  Taken
+together, these constitute a score by chaining together multiple pairs
+over the course of time:
 
@@
 
-%% \[
-%%     f : image^{\ast} \to (
-%% \]
+However, producing a MIDI score with only a single instrument would
+sound broing (akin to a song with only one player): instead we take
+pieces of images and map them into different parts.  Each of these
+parts constitutes an instrument and --- when mixed together ---
+produces a final score.  An illustration of the core idea is shown in
+Figure~\ref{fig:images-to-music}.
+
+\begin{figure*}
+  \centering
+  \begin{minipage}[t]{.3\textwidth}
+    \centering
+    \includegraphics[width=.7\textwidth]{images-to-regions.png}
+  \end{minipage}
+  \begin{minipage}[t]{.1\textwidth}
+    \raisebox{.8 in}{\Huge $\to$}
+  \end{minipage}
+  \begin{minipage}[t]{.4\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{parts-and-instruments.png}
+  \end{minipage}
+  \caption{Mapping a stream of three images to a nine piece musical
+    score.  First each image is mapped into regions, which will be
+    transformed into notes and combined into a score using a mixer.}
+\end{figure*}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.4\textwidth]{hsv.png}
+  \caption{The HSV color cylinder}
+  \label{fig:hsv}
+\end{figure}
 
@@
 saturation, and a value.  These components are visualized (as shown in
-Figure~\ref{fig:hsv}) using a wheel.  The components represent:
+Figure~\ref{fig:hsv}) as a cylinder (as opposed to RGB, which views
+the color space as a cube).  Each component represents:
 
 \begin{enumerate}
-\item Hue --- The actual ``color'' that would be described.  At full
+\item Hue: The actual ``color'' that would be described.  At full
   saturation and value, the hue determines the color seen.  This is
@@
 
-\item Saturation --- The degree to which the color is saturated from
+\item Saturation: The degree to which the color is saturated from
   white.  Zero saturation is a blank (white) surface, while increasing
@@
 
-\item Value --- The amount of light in the color.  Colors with low
-  values are dark, and those with higher values are brighter.  This is
+\item Value: The amount of light in the color.  Colors with low values
+  are dark, and those with higher values are brighter.  This is
   visualized by going up the cylinder.
@@
 
-\kris{Here I'm going to draw a diagram which shows two points on the
-  HSV wheel and a musical clef, illustrating the position.}
-
 In our transformation to music, we map colors in HSV to notes in MIDI.
-To do this, we map colors in RGB to HSV using the \code{\#RGBtoHSV}
-method of the \code{java.awt.Color} class.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.4\textwidth]{hsv.png}
-  \caption{The HSV color cylinder}
-  \label{fig:hsv}
-\end{figure}
+We do this by overlaying the circle of fifths (a circle representing
+pleasing intervals called fifths in music theory) onto the circle
+representing hue in HSV.  We establish the octave of the produced note
+using the value of the color in HSV, and the volume of the note using
+the saturation.
 
@@
 
-\begin{figure}
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.4\linewidth]{image1}
-  \caption{A subfigure}
-  \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.4\linewidth]{image1}
-  \caption{A subfigure}
-  \label{fig:sub2}
-\end{subfigure}
-\caption{A figure with two subfigures}
-\label{fig:test}
-\end{figure}
-
 \section{Implementation}
@@
 MapReduce phases, and a music mixing tool.  The high level pipeline is
-shown in Figure~\ref{fig:project-pipeline}.
+shown in Figure~\ref{fig:project-pipeline}.  It consists of a sequence
+of MapReduce jobs, along with sets of helper scripts
 
-\paragraph{Conversion of images to thumbnails}
+\subsection{Conversion of images to thumbnails}
 
@@
 
-\paragraph{Conversion to intermediate representation}
+\subsection{Conversion to intermediate representation}
 \label{sec:conversion}
@@
 \begin{lstlisting}[language=Java]
-public class WriteImagesToSequenceFile {
+public class WriteImagesToSequenceFile
+{
   BufferedImage img = null;
   SequenceFile.Writer writer = null;
-  FileSystem fs = FileSystem.get(conf);
-
+  FileSystem fs = 
+      FileSystem.get(conf);
   FileStatus[] fss = 
-    fs.globStatus(new Path(inputPath 
+    fs.globStatus(new Path(input 
       + "/*.jpg"));
-
-  writer = SequenceFile.createWriter(fs, 
-    conf, outPath, IntWritable.class,
+  writer = SequenceFile
+      .createWriter(fs, 
+    conf,outPath,IntWritable.class,
     ArrayListOfIntsWritable.class);
@@
         new ArrayListOfIntsWritable();
-
     img = ImageIO.read(
@@
            .toString().substring(5)));
-
-    // Get the RBG integer for each pixel 
-    // in 3X3
+    // Get RBG integer for each pixel 
     for(int j = 0; j < thumbs; j++) {
       for(int i = 0; i< thumbs; i++) {
-        pixelData.add(img.getRGB(i,j));
+        pixelData
+          .add(img.getRGB(i,j));
       }
     }
-
     // Write to the sequence file
     image_counter.set(counter);
-    writer.append(image_counter, pixelData);
+    writer.append(image_counter,
+      pixelData);
     counter += 1;
@@
 
-\paragraph{Translating to note, region pairs}
+\subsection{Translating to note, region pairs}
 
@@
 collate together.  After mapping over each pixel, we need to end up
-with a stream of images for each region: a naive implementation simply
+with a stream of images for each region: a na\"{i}ve implementation simply
 maps each pixel in our input corpus to these tone, velocity pairs:
@@
 \]
+
+A snippet of our mapper is shown in
+Figure~\ref{figure:colortomusic-mapper}.
+
+\begin{figure}
+\label{figure:colortomusic-mapper}
+\begin{lstlisting}[language=Java]
+public void map(IntWritable key, 
+    ArrayListOfIntsWritable pixels,
+    Context context)
+  throws IOException, InterruptedException {
+  int regionCounter = 1;
+  MidiNote midiNote = null;
+  for(int i = 0; i < pixels.size(); i++){
+    pixel = pixels.get(i);
+    // Conversion from color to note
+    midiNote = Color2Music
+      .convert(new Color(pixel));
+    REGION_IMAGENO.set(regionCounter,
+      key.get());
+    NOTE_VELOCITY.set(midiNote.getTone(),
+      midiNote.getVelocity());
+    context.write(REGION_IMAGENO, NOTE_VELOCITY);
+    regionCounter++;
+  }
+}
+\end{lstlisting}
+\caption{A portion of our mapper that takes color data to MIDI notes.
+  For each pixel in the original image, the conversion in
+  \code{Color2Music} is applied before setting the output key (as a
+  region, number pair).}
+\end{figure}
 

### 2013-05-10 writeup/paper.tex
adding stuff to the conclusion

@@
 \begin{lstlisting}[language=Java]
-public class WriteImagesToSequenceFile
+public class WriteImgsToSequence
 {
   BufferedImage img = null;
-  SequenceFile.Writer writer = null;
+  SequenceFile.Writer writer=null;
   FileSystem fs = 
@@
          .toString().substring(5)));
-    // Get RBG integer for each pixel 
-    for(int j = 0; j < thumbs; j++) {
-      for(int i = 0; i< thumbs; i++) {
+    // Get RBG for each pixel 
+    for(int j = 0; j<thumbs; j++) {
+     for(int i = 0; i<thumbs; i++) {
         pixelData
@@
     Context context)
-  throws IOException, InterruptedException {
+{
   int regionCounter = 1;
   MidiNote midiNote = null;
-  for(int i = 0; i < pixels.size(); i++){
+  for(int i=0; i<pixels.size();i++){
     pixel = pixels.get(i);
@@
       midiNote.getVelocity());
-    context.write(REGION_IMAGENO, NOTE_VELOCITY);
+    context.write(REGION_IMAGENO,
+       NOTE_VELOCITY);
     regionCounter++;
@@
 
-The result of our mapreduce pipeline is a set of MIDI music: one for
-each region in the image set.  To turn these into music that can be
-played, we use mixing software (such as GarageBand) to mix the regions
-into full scores.
+The result of our mapreduce pipeline is a set of notes in MIDI: one
+for each region in the image set.  We postprocess the MapReduce output
+by taking it off HDFS using a script and converting to the MIDI format
+(adding on a header about the number of parts and other small details
+to be a valid MIDI file). To turn these into music that can be played,
+we use mixing software (such as GarageBand) to mix the regions into
+full scores and generate MP3 files.
+
+To visualize music and colors, we take the resulting music and mix it
+with a video generated from the thumbnailed images in our input
+dataset.  We do this using a process similar to preprocessing: we
+simply blow up the thumbnailed images to a larger size, concatenate
+them together into subsequent frames of a video, and mix the video
+with the music generated from our MapReduce phases.
 
@@
 
-\subsection{Sample datasets}
-
 To test the system, we use a variety of different sets of images.
 These images were obtained from
-Flickr\footnote{\url{http://www.flickr.com/}}, using a public API and
-a Python script \cite{flickerpy} to get test data.
+Flickr\footnote{\url{http://www.flickr.com/}} and \code{memebase.com},
+using a public API and a Python script \cite{flickerpy} to get test
+data.  We converted animated GIFs into individual frames again using a
+preprocessing phase.  Our MapReduce code takes as input simply a
+\code{SequenceFile} (as discussed in Section~\ref{sec:conversion}), so
+we simply feed in a folder of JPG images to our preprocessing phase.
+
+As a result, we produced various pieces of music.  Our longest piece
+(named \code{Digital Nightmare}) was over six minutes long and
+composed of over 3000 images.  We also produced shorter (self
+contained) pieces from various related images: such as GIFs of cat
+jumping animations.  All of these pieces are made available in our
+repository.
+
+Our MapReduce phases ran very quickly (on the order of minutes, mostly
+startup time).  We believe that the reason for this is that by the
+time preprocessing was completed we ended up with a relatively small
+amount of data to be fed into MapReduce.  Were we to do the project
+over again we would have fed in much more data to the MapReduce jobs
+and optimized the selection of regions (from each image) to be the
+``best sounding'' piece of music that could be produced.  We thought
+about how to devise a measure of ``realistic'' music, but could not
+easily see one.  One potential next application would be to use Hadoop
+to mine this correlation.  The idea is that realistic music generally
+has some recurring nature (backbeat, for example) even if very
+complex.  Therefore, we would want to apply a penalty term to our
+measure whenever we generated music that had too complex a principal
+structure (this would avoid creating chaotic sounding music).  We did
+some initial work on this (writing down some equations, studying how
+to analyze MIDI for it's beat) but didn't have anything concrete by
+the time we needed to do implementation.
+
+We also noticed that there might be some interesting problems in
+taking a song and a set of videos and finding out which best matched
+each other.  (In essence, this would let you search for image based on
+video.)  We started to write an inverse transformation within
+MapReduce\footnote{src/main/phase2/Color2Music} to explore this idea,
+but haven't finished doing the work yet.
 

### 2013-05-10 src/main/phase1/JpgToBitmapPreprocessing.java
Removing files not needed for submission

@@ -1,130 +1,0 @@
-package phase1;
-
-import java.io.IOException;
-import java.util.Arrays;
-
-import org.apache.commons.cli.CommandLine;
-import org.apache.commons.cli.CommandLineParser;
-import org.apache.commons.cli.GnuParser;
-import org.apache.commons.cli.HelpFormatter;
-import org.apache.commons.cli.OptionBuilder;
-import org.apache.commons.cli.Options;
-import org.apache.commons.cli.ParseException;
-import org.apache.hadoop.conf.Configuration;
-import org.apache.hadoop.conf.Configured;
-import org.apache.hadoop.fs.FileSystem;
-import org.apache.hadoop.fs.Path;
-import org.apache.hadoop.io.IntWritable;
-import org.apache.hadoop.mapreduce.Job;
-import org.apache.hadoop.mapreduce.Mapper;
-import org.apache.hadoop.mapreduce.lib.input.FileInputFormat;
-import org.apache.hadoop.mapreduce.lib.input.SequenceFileInputFormat;
-import org.apache.hadoop.mapreduce.lib.output.FileOutputFormat;
-import org.apache.hadoop.mapreduce.lib.output.SequenceFileOutputFormat;
-import org.apache.hadoop.util.Tool;
-import org.apache.hadoop.util.ToolRunner;
-import org.apache.log4j.Logger;
-
-public class JpgToBitmapPreprocessing extends Configured implements Tool {
-  private static final Logger LOG = Logger.getLogger(JpgToBitmapPreprocessing.class);
-
-  
-  // Mapper: emits ...
-  private static class ImageToIntermediateMapper extends Mapper<IntWritable, IntWritable, IntWritable, IntWritable>{
-
-
-    public void map(IntWritable key, IntWritable value, Context context)
-      throws IOException, InterruptedException {
-      
-      // Read in image data and create intermediate image data representation
-      
-    }
-  }
-  
-  
-  public JpgToBitmapPreprocessing() {
-  }
-  
-  private static final String INPUT = "input";
-  private static final String OUTPUT = "output";
-  
-  /**
-   * Runs this tool.
-   */
-  @SuppressWarnings("static-access")
-  @Override
-  public int run(String[] args) throws Exception {
-    
-    Options options = new Options();
-    
-    options.addOption(OptionBuilder.withArgName("path").hasArg()
-        .withDescription("input path").create(INPUT));
[…]

### 2013-05-10 src/main/phase2/GrabRGBFromImage.java
Removing file used to test approach

@@ -1,66 +1,0 @@
-package phase2;
-
-import java.awt.Color;
-import java.awt.image.BufferedImage;
-import java.io.ByteArrayInputStream;
-import java.io.ByteArrayOutputStream;
-import java.io.File;
-import java.io.IOException;
-import java.io.ObjectInputStream;
-import java.io.ObjectOutput;
-import java.io.ObjectOutputStream;
-
-import javax.imageio.ImageIO;
-
-import org.apache.hadoop.io.BytesWritable;
-
-public class GrabRGBFromImage {
-
-  public static void main(String[] args) throws IOException {
-    BufferedImage img = null;
-    ByteArrayOutputStream bOut = new ByteArrayOutputStream();
-    ObjectOutput out = new ObjectOutputStream(bOut);
-
-    // Read an image from file
-    try {
-      img = ImageIO.read(new File("images/collection1-thumbs/thumb-sample-01.jpg"));
-    } catch (IOException e) {
-    
-    }
-
-    // Get the RGB integer for each pixel in 3X3
-    for(int j = 0; j < 3; j++) {
-      for(int i = 0; i < 3; i++) {
-        System.out.println("Reading pixel: i = " + i + ", j = " + j);
-        out.writeInt(img.getRGB(i, j));
-      }
-    }
-    out.close();
-
-    // Make a BytesWritable and with the output stream data
-    BytesWritable BW = new BytesWritable(bOut.toByteArray());
-    bOut.close();
-
-
-    byte[] imgBytes = BW.getBytes();
-    ByteArrayInputStream inStream = new ByteArrayInputStream(imgBytes);
-
-    ObjectInputStream ois = new ObjectInputStream(inStream);
-
-    int pixel = -1;
-    Color c = null;
-    while(ois.available() > 0){
-      pixel = ois.readInt();
-      c = new Color(pixel);
-     
-      System.out.print("R: " + c.getRed() + "\t");
-      System.out.print("G: " +  c.getGreen() + "\t");
-      System.out.print("B: "+  c.getBlue());
-      System.out.println();
-    }
-
-    ois.close();
-
-  }
-
-}

### 2013-05-10 src/main/phase2/Color2Music.java
Cleaning up

@@ -1,9 +1,5 @@
 package phase2;
 
-import java.awt.Color;
-
-
-
 public class Color2Music {
 
   public static final int LOWEST_NOTE = 36;
@@ -66,8 +62,6 @@ public class Color2Music {
   }
   
   public static MidiNote convert(int pixel){
-
-    
     RGBtoHSV(pixel);
     
     float hue = hsvConversionArray[0];
@@ -75,13 +69,6 @@ public class Color2Music {
     float brightness = hsvConversionArray[2];
     
     int tone = getMidiToneFromHSValues(hue, saturation, brightness);
-    
-//    
-//    int velocity = saturationFloatToVelocityInt(saturation);
-//    if(velocity > MIN_VELOCITY_THRESHOLD){
-//      tone = getMidiToneFromHSValues(hue, saturation, brightness);
-//    }
-
     return new MidiNote(tone, 127);
   }
   
@@ -94,89 +81,23 @@ public class Color2Music {
   
   private static int brightessAndSaturationToOctaveInt(float s, float b){
     
-    
-    System.out.print("b: " + b + "\ts: " + s );
-    
     double bWeight = 0.5;
     double sWeight = 1 - bWeight;
 
     double maxSContribution = sWeight * OCTAVE_RANGE;
     double maxBContribution = bWeight * OCTAVE_RANGE;
     
-//    System.out.println("- maxsat: " + maxSContribution);
-//    System.out.println("- maxbri: " + maxBContribution);
-    
-    
-    
-    // faux "absolute value"
-    // s below 0.5 maps to (1 - s) (ie put it in the 0.5 -> 1.0 range
     double satAbs = s < 0.5 ? (1 - s) : s;
     double sValue = Math.floor(satAbs / (1.0 / maxSContribution));
-
- // Number between 0 and (maxBContribution - 1)
     double bValue = Math.floor(b / (1.0 / maxBContribution));
-    
-//    System.out.println(" --- sValue: " + sValue);
-//    System.out.println(" --- bValue: " + bValue);
     
     int octave = (int) Math.floor(bValue + sValue);
     
-    
-    
-//    if(b < 0.2) octave = 0;
-    
-    
-    System.out.println(" -- \tOctave: " + octave );
-//    System.out.println();
     return octave;
-    
-//    int brightnessComponent = (int) Math.floor(b / (1.0 / OCTAVE_RANGE));
-//    int satOn = 0;
-//    
-//    if(s < 0.33 || s > 0.67 ){
-//      satOn = 1;
[…]

### 2013-05-10 writeup/paper.tex
Added abstract

@@
 \begin{abstract}
+Every young computer scientist learns early in their career that all data regardless of its meaning are all the same in their underlying representation - simple ones and zeroes.
+It is only in the act of reading and subsequently interpreting those data that computer scientists or their creations give meaning to the bits and bytes.
+MapReduce to My Ears is an exercise in the reinterpretation of data meant to be understood as color and picture into sound and music.
+Our method uses Hadoop, an open source implementation of the MapReduce programming paradigm, to process a large corpus of images and extract a musical representation.
+We have produce a six minutes and thirty-five second audio/visual entitled \emph{Digital Nightmare}, the style of which can be described as 8-bit, blast beat.
 \end{abstract}

### 2013-05-10 writeup/paper.tex
adding amanda's changes

--- 2013-05-10 writeup/paper.tex
@@
 \begin{abstract}
+
 \end{abstract}
@@
 different areas of computer science have been lifted into the
-MapReduce framework.  By contrast, MapReduce has been used relatively
-less to explore art or creative tasks.  Our project looks to answer:
-what kinds of art can we create harnessing large amounts of data at
-scale?
+MapReduce framework.  While MapReduce has been applied to many
+large-scale data processing domains, it has been used relatively less
+to explore art or creative tasks.  Our project looks to answer: what
+kinds of art can we create harnessing large amounts of data at scale?
 
@@
 system we chose MIDI as the output format.  Our system can then be
-modeled as a function from image data to MIDI output.  Note that the
-choice of MIDI limits the expressivity of our music.  Going to other
-formats (such as MP3 or OOG) could provide a larger range of
-primitives with which to construct music, but does not enjoy the
-simplicity that MIDI provides.
+modeled as a function from image data to MIDI output.  Using MIDI as a
+format works well because it is easy to work with (we can model music
+as MIDI events, which are simply streams of musical data) and can be
+mixed by many synthesizers to produce music using a variety of
+instruments.
 
@@
 (named \code{Digital Nightmare}) was over six minutes long and
-composed of over 3000 images.  We also produced shorter (self
+composed of over 3500 images.  We also produced shorter (self
 contained) pieces from various related images: such as GIFs of cat
 jumping animations.  All of these pieces are made available in our
-repository.
+repository.  We ran the entire MapReduce job on the Hoth cluster to
+produce our longest piece.
 
@@
 
-Many domain specific programming languages exist for music synthesis
-[\cite{haskore},\cite{lilypad}].  Many of these languages synthesize
-music in the small ().  By contrast, our approach focuses on using
-large amounts of data to create music...
+Our work is inspired by the music visualization effects included in
+media player software (which traditionally includes effects such as
+oscilloscopes and frequency representations).  However, many domain
+specific programming languages exist for music synthesis.  In
+particular we looked at Haskore \cite{haskore} and
+LilyPond\cite{lilypond}.  Haskore is a language based on Haskell,
+which allows the user to express mathematically beautiful music using
+an algebraic composition strategy.  LilyPond is a specification
+language used to create high quality sheet music (similar to \LaTeX\ 
+for music).  Many of these languages synthesize music in the small,
+using a human based specification language.  By contrast, our approach
+focuses on using large amounts of data to create music: rather than a
+human specification.  We are unaware of any projects that use big data
+to produce art or other forms of entertainment.
 

### 2013-05-10 writeup/paper.tex

@@
 \begin{abstract}
-
+Every young computer scientist learns early in their career that all
+data regardless of its meaning are all the same in their underlying
+representation - simple ones and zeroes.
+ It is only in the act of
+reading and subsequently interpreting those data that computer
+scientists or their creations give meaning to the bits and bytes.
+MapReduce to My Ears is an exercise in the reinterpretation of data
+meant to be understood as color and picture into sound and music.
+Our method uses Hadoop, an open source implementation of the MapReduce
+programming paradigm, to process a large corpus of images and extract
+a musical representation.
+ We have produce a six minutes and
+thirty-five second audio/visual entitled \emph{Digital Nightmare}, the
+style of which can be described as 8-bit, blast beat.
 \end{abstract}

### 2013-05-10 writeup/paper.tex
Added reflections subsection

--- 2013-05-10 writeup/paper.tex
@@
 
+\subsection{Reflection on Results}
+
+One might describe our final piece, \emph{Digital Nightmare}, as an unrelenting flood of almost patternless musical information.
+This characteristic emerges from some simplifications in design choices that a later vernon of our process would look to address.
+Here we present a list of such areas and possible solutions.
+\begin{itemized}
+\item Each instrument plays a note during every musical time unit (beat).
+This results in constant note information without any space for the listener to "catch up".
+Darker colors below some threshold of brightness could all be translated into musical rests.
+\item Each note played last only one beat, i.e quarter notes are the only length note used.
+Consecutive image regions with similar colors could be "strung together" cause some instruments to play longer notes.
+\item Late on in the process, we moved from saturation mapping to velocity to choosing a fixed velocity of 127.
+We found that velocity dynamics were not so perceptible with the degree of musical production we were willing to invest.
+Spending more time in the post-processing phase and focusing on musical dynamics such as volume, panning, etc could improve the contribution of velocity to a musical piece.
+\item In out final piece, saturation did not contribute to velocity, but instead was a component of the octave calculation.
+This choice was made after observing some short musical pieces and feeling that the color to sound mapping was not as intuitive as we had expected.
+This change helped somewhat, but we feel the issue was related to a number of factors including the limitations of the HSB color model to represent the human visual model, the discretization of a continuous range (color is continuous whereas music notes are discrete), and using a constant scale rather than a logarithmic scale to represent musical frequencies (which are inherently logarithmic, each an octave higher is twice the lower octave's frequency).
+\end{itemized}
+
 \section{Conclusion}

### 2013-05-10 writeup/paper.tex
removed code examples, fixed things amanda mentioned

--- 2013-05-10 writeup/paper.tex
@@
 Our system takes in ordered sets of images and produces a MIDI file.
-In our view, a MIDI file consists of note and duration pairs.  Taken
-together, these constitute a score by chaining together multiple pairs
-over the course of time:
+The input format is JPEG images: this is a very commonly used lossy
+compression format.  Since we used input corpuses from the internet,
+using JPEG worked for the variety of input formats.  Using JPEG as the
+input format is also attractive because it can be easily unpacked by
+the Java \code{ImageBuffer} class.
+
+The output of our MapReduce phases is MIDI music.  MIDI is a simple
+format for writing music which can be mixed ans synthesized using
+software.  At an abstract level, a MIDI file consists of note and
+duration pairs.  Taken together, these constitute a score by chaining
+together multiple pairs over the course of time:
 
@@
 \label{sec:conversion}
-
 Image data comes in a variety of different formats: JPEG, GIF, PNG,
@@
 Section~\ref{sec:overview}) the first stage of our Hadoop
-implementation\footnote{\code{src/main/phase1}} converts the surface
-image representation to the sequence file format used in subsequent
-processing.  Although it may appear inefficient to hold an
-uncompressed image in memory, this is actually ideal for Hadoop:
-because computing cycles (incurred through recompressing and sending
-the images) are the cost, compared with network traffic.
+implementation converts the surface image representation to the
+sequence file format used in subsequent processing.  Although it may
+appear inefficient to hold an uncompressed image in memory, this is
+actually ideal for Hadoop: because computing cycles (incurred through
+recompressing and sending the images) are the cost, compared with
+network traffic.
 
-The conversion in our prototype uses the java \code{ImageBuffer} class
-to do image input and conversion.  This class holds requires a large
-amount of memory to represent image objects, however this is fine
-because we hold only a small number of images in memory at any one
-time (using statics) when mapping over our input corpus.
+The conversion in our prototype uses the java \code{ImageBuffer}
+class to do image input and conversion.  This class holds requires a
+large amount of memory to represent image objects, however this is
+fine because we hold only a small number of images in memory at any
+one time (using statics) when mapping over our input corpus.
 
@@
 
-\begin{figure}
-\begin{lstlisting}[language=Java]
-public class WriteImgsToSequence
-{
-  BufferedImage img = null;
-  SequenceFile.Writer writer=null;
-  FileSystem fs = 
-      FileSystem.get(conf);
-  FileStatus[] fss = 
-    fs.globStatus(new Path(input 
-      + "/*.jpg"));
-  writer = SequenceFile
-      .createWriter(fs, 
-    conf,outPath,IntWritable.class,
-    ArrayListOfIntsWritable.class);
-  int counter = 0;
-  for (FileStatus status : fss) {
-    IntWritable image_counter = 
-      new IntWritable();
-    ArrayListOfIntsWritable 
-     pixelData = 
-      new ArrayListOfIntsWritable();
-    img = ImageIO.read(
-      new File(status.getPath()
-         .toString().substring(5)));
-    // Get RBG for each pixel 
-    for(int j = 0; j<thumbs; j++) {
-     for(int i = 0; i<thumbs; i++) {
-        pixelData
-          .add(img.getRGB(i,j));
-      }
-    }
-    // Write to the sequence file
-    image_counter.set(counter);
-    writer.append(image_counter,
-      pixelData);
-    counter += 1;
-  } 
-}
-\end{lstlisting}
-\caption{Snippet of code from \code{WriteToSequenceFile} showing
-  format used in sequence files throughout the MapReduce jobs.}
-\end{figure}
-
 \subsection{Translating to note, region pairs}
@@
 portions of each image.  The conversion happens by mapping over our
-set of images in the \code{Image2MusicMR} class.  The output of the
-image to music conversion is $n$ sets of note lists (where notes are
-expressed as pairs).
+set of images in the \code{Image2MusicMR} class.  The input of this
+reducer is a set of key/value pairs which constitute the image number
+and image data (a set of integers representing each pixel in the
+image).  The output of the image to music conversion is $n$ sets of
+note lists (where notes are expressed as pairs).  This is stored as a
+set of key/value pairs where the key is the region number and image
+number pair (subsequently discussed, but used for secondary sorting)
+and the part number and MIDI note.
 
@@
 \]
-
-A snippet of our mapper is shown in
-Figure~\ref{figure:colortomusic-mapper}.
-
-\begin{figure}
-\label{figure:colortomusic-mapper}
-\begin{lstlisting}[language=Java]
-public void map(IntWritable key, 
-    ArrayListOfIntsWritable pixels,
-    Context context)
-{
-  int regionCounter = 1;
-  MidiNote midiNote = null;
-  for(int i=0; i<pixels.size();i++){
-    pixel = pixels.get(i);
-    // Conversion from color to note
-    midiNote = Color2Music
-      .convert(new Color(pixel));
-    REGION_IMAGENO.set(regionCounter,
-      key.get());
-    NOTE_VELOCITY.set(midiNote.getTone(),
-      midiNote.getVelocity());
-    context.write(REGION_IMAGENO,
-       NOTE_VELOCITY);
-    regionCounter++;
-  }
-}
-\end{lstlisting}
-\caption{A portion of our mapper that takes color data to MIDI notes.
-  For each pixel in the original image, the conversion in
-  \code{Color2Music} is applied before setting the output key (as a
-  region, number pair).}
-\end{figure}
 

### 2013-05-10 writeup/paper.tex
removed more code

@@
 
-\begin{figure}[h]
+\begin{figure}
   \centering
@@
 \cite{imagemagick}.  ImageMagick includes facilities for taking images
-to thumbnails.
+to thumbnails, we simply use the RMagick library to preprocess all of
+our images so that they are in the correct dimensions and are of the
+proper size.
 
-\begin{figure}
-\begin{lstlisting}[language=Ruby]
-require 'rmagick'
+%% \begin{figure}
+%% \begin{lstlisting}[language=Ruby]
+%% require 'rmagick'
 
-class ThumbMaker
-  # Convert set of images to thumbnails
-  def self.convert_thumbs(dir_path, size)
-      # ... 
-      puts "Making thumbnail for..."
-      Dir.glob("#{path}/*.jpg") do |fname|
-        # ...
-        fname = # ...
-        img = Magick::Image.read(fname)[0]
-        thumb = img.thumbnail(size, size)
-        thumb.write(thumb_dest)
-      end
-    end
-end
+%% class ThumbMaker
+%%   # Convert set of images to thumbnails
+%%   def self.convert_thumbs(dir_path, size)
+%%       # ... 
+%%       puts "Making thumbnail for..."
+%%       Dir.glob("#{path}/*.jpg") do |fname|
+%%         # ...
+%%         fname = # ...
+%%         img = Magick::Image.read(fname)[0]
+%%         thumb = img.thumbnail(size, size)
+%%         thumb.write(thumb_dest)
+%%       end
+%%     end
+%% end
 
-\end{lstlisting}
-\caption{Relevant code from the module used for image to thumbnail
-  conversion}
-\end{figure}
+%% \end{lstlisting}
+%% \caption{Relevant code from the module used for image to thumbnail
+%%   conversion}
+%% \end{figure}
 

### 2013-05-10 writeup/paper.tex
adding makefile, citations, and cleanups

@@
 Flickr\footnote{\url{http://www.flickr.com/}} and \code{memebase.com},
-using a public API and a Python script \cite{flickerpy} to get test
+using a public API and a Python script \cite{flickrpy} to get test
 data.  We converted animated GIFs into individual frames again using a

### 2013-05-10 writeup/paper.tex
Adding conclusion

@@
 
+At the onset of this project, we had a vision for translating images in music and we have succeeded in doing so.
+Just about every design decision in our process has room for improvement, we are pleased to have at least found a path from start to finish, though not without trials.
+At every stage of the pipeline, there were new complexities to discover - the various color models and their relationships, the difficulty in using natural language to discuss color, or the vast number of parameters needed for machines to represent a piece of audio or visual art.
+However challenging though, the journey has put into perspective the amazing ability of the painter and composer to produce works of art with far more complexity than machines can manage. 
+We hope that our work inspires technologists to apply their skills to reinterpret works of others from unrelated fields, and vice versa.
+
 \bibliographystyle{IEEEtran}

### 2013-05-10 writeup/paper.tex
Readded reflections, might have gotten clobbered

@@
 
+\subsection{Reflection on Results}
+
+One might describe our final piece, \emph{Digital Nightmare}, as an unrelenting flood of almost patternless musical information.
+This characteristic emerges from some simplifications in design choices that a later vernon of our process would look to address.
+Here we present a list of such areas and possible solutions.
+\begin{itemized}
+\item Each instrument plays a note during every musical time unit (beat).
+This results in constant note information without any space for the listener to "catch up".
+Darker colors below some threshold of brightness could all be translated into musical rests.
+\item Each note played last only one beat, i.e quarter notes are the only length note used.
+Consecutive image regions with similar colors could be "strung together" cause some instruments to play longer notes.
+\item Late on in the process, we moved from saturation mapping to velocity to choosing a fixed velocity of 127.
+We found that velocity dynamics were not so perceptible with the degree of musical production we were willing to invest.
+Spending more time in the post-processing phase and focusing on musical dynamics such as volume, panning, etc could improve the contribution of velocity to a musical piece.
+\item In out final piece, saturation did not contribute to velocity, but instead was a component of the octave calculation.
+This choice was made after observing some short musical pieces and feeling that the color to sound mapping was not as intuitive as we had expected.
+This change helped somewhat, but we feel the issue was related to a number of factors including the limitations of the HSB color model to represent the human visual model, the discretization of a continuous range (color is continuous whereas music notes are discrete), and using a constant scale rather than a logarithmic scale to represent musical frequencies (which are inherently logarithmic, each an octave higher is twice the lower octave's frequency).
+\end{itemized}
+
 \section{Related work}

### 2013-05-10 writeup/paper.tex
small changes.

--- 2013-05-10 writeup/paper.tex
@@
 
-The big data revolution has revolutionized the way we think about,
-design, and implement algorithms at scale.  As such, problems from all
-different areas of computer science have been lifted into the
-MapReduce framework.  While MapReduce has been applied to many
-large-scale data processing domains, it has been used relatively less
-to explore art or creative tasks.  Our project looks to answer: what
-kinds of art can we create harnessing large amounts of data at scale?
+Big data has revolutionized the way we think about, design, and
+implement algorithms at scale.  As such, problems from all different
+areas of computer science have been lifted into the MapReduce
+framework.  While MapReduce has been applied to many large-scale data
+processing domains, it has been used relatively less to explore art or
+creative tasks.  Our project looks to answer: what kinds of art can we
+create harnessing large amounts of data at scale?
 
@@
 The input is a large corpus of images, which is subsequently processed
-by our system, and the result is a musical score.  We then mix and
-listen to the music.
+by our system, and the result is a musical score.  We then mix,
+listen, and refine the algorithm to produce art.
 

### 2013-05-10 writeup/paper.tex
fixing TeX

@@
 Here we present a list of such areas and possible solutions.
-\begin{itemized}
-\item Each instrument plays a note during every musical time unit (beat).
-This results in constant note information without any space for the listener to "catch up".
-Darker colors below some threshold of brightness could all be translated into musical rests.
-\item Each note played last only one beat, i.e quarter notes are the only length note used.
-Consecutive image regions with similar colors could be "strung together" cause some instruments to play longer notes.
-\item Late on in the process, we moved from saturation mapping to velocity to choosing a fixed velocity of 127.
-We found that velocity dynamics were not so perceptible with the degree of musical production we were willing to invest.
-Spending more time in the post-processing phase and focusing on musical dynamics such as volume, panning, etc could improve the contribution of velocity to a musical piece.
-\item In out final piece, saturation did not contribute to velocity, but instead was a component of the octave calculation.
-This choice was made after observing some short musical pieces and feeling that the color to sound mapping was not as intuitive as we had expected.
-This change helped somewhat, but we feel the issue was related to a number of factors including the limitations of the HSB color model to represent the human visual model, the discretization of a continuous range (color is continuous whereas music notes are discrete), and using a constant scale rather than a logarithmic scale to represent musical frequencies (which are inherently logarithmic, each an octave higher is twice the lower octave's frequency).
-\end{itemized}
+
+\begin{itemize}
+
+\item Each instrument plays a note during every musical time unit
+  (beat). This results in constant note information without any space
+  for the listener to "catch up".  Darker colors below some threshold
+  of brightness could all be translated into musical rests.
+
+\item Each note played last only one beat, i.e quarter notes are the
+  only length note used.  Consecutive image regions with similar
+  colors could be "strung together" cause some instruments to play
+  longer notes.
+
+\item Late on in the process, we moved from saturation mapping to
+  velocity to choosing a fixed velocity of 127.  We found that
+  velocity dynamics were not so perceptible with the degree of musical
+  production we were willing to invest.  Spending more time in the
+  post-processing phase and focusing on musical dynamics such as
+  volume, panning, etc could improve the contribution of velocity to a
+  musical piece.
+
+\item In our final piece, saturation did not contribute to velocity,
+  but instead was a component of the octave calculation. This choice
+  was made after observing some short musical pieces and feeling that
+  the color to sound mapping was not as intuitive as we had expected.
+  This change helped somewhat, but we feel the issue was related to a
+  number of factors including the limitations of the HSB color model
+  to represent the human visual model, the discretization of a
+  continuous range (color is continuous whereas music notes are
+  discrete), and using a constant scale rather than a logarithmic
+  scale to represent musical frequencies (which are inherently
+  logarithmic, each an octave higher is twice the lower octave's
+  frequency).
+
+\end{itemize}
 
@@
 
-At the onset of this project, we had a vision for translating images in music and we have succeeded in doing so.
-Just about every design decision in our process has room for improvement, we are pleased to have at least found a path from start to finish, though not without trials.
-At every stage of the pipeline, there were new complexities to discover - the various color models and their relationships, the difficulty in using natural language to discuss color, or the vast number of parameters needed for machines to represent a piece of audio or visual art.
-However challenging though, the journey has put into perspective the amazing ability of the painter and composer to produce works of art with far more complexity than machines can manage. 
-We hope that our work inspires technologists to apply their skills to reinterpret works of others from unrelated fields, and vice versa.
+At the onset of this project, we had a vision for translating images
+in music and we have succeeded in doing so.  Just about every design
+decision in our process has room for improvement, we are pleased to
+have at least found a path from start to finish, though not without
+trials.  At every stage of the pipeline, there were new complexities
+to discover --- the various color models and their relationships, the
+difficulty in using natural language to discuss color, or the vast
+number of parameters needed for machines to represent a piece of audio
+or visual art.  However challenging though, the journey has put into
+perspective the amazing ability of the painter and composer to produce
+works of art with far more complexity than machines can manage.  We
+hope that our work inspires technologists to apply their skills to
+reinterpret works of others from unrelated fields, and vice versa.
 

### 2013-05-10 writeup/paper.tex
Fixed typo
small changes to things

@@
 representing, etc..).  We anticipate that exploring this area might
-open up potential applications.
+open up potential applications involving creation of art using amounts
+of data never before considered possible.
 
@@
 compression format.  Since we used input corpuses from the internet,
-using JPEG worked for the variety of input formats.  Using JPEG as the
-input format is also attractive because it can be easily unpacked by
-the Java \code{ImageBuffer} class.
+many of the publicly available image sets are natively distributed in
+JPEG.  Using JPEG as the input format is also attractive because it
+can be easily unpacked by the Java \code{ImageBuffer} class, which we
+use throughout our implementation.
 

### 2013-05-10 writeup/paper.tex
small fixes to refs
introducing HSV more clearly

@@
     transformed into notes and combined into a score using a mixer.}
+  \label{fig:images-to-music}
 \end{figure*}
@@
 bijection between the data in the image and the range of music in the
-song.  The way we do this is inspired by viewing colors in images as
-HSV values.  In HSV, each color has three components: a hue, a
-saturation, and a value.  These components are visualized (as shown in
+song.  The way we do this is inspired by viewing colors in images
+using a slightly unconventional color representation system: HSV.  In
+HSV, each color has three components: a hue, a saturation, and a
+value.  These components are visualized (as shown in
 Figure~\ref{fig:hsv}) as a cylinder (as opposed to RGB, which views
@@
 
-In our transformation to music, we map colors in HSV to notes in MIDI.
-We do this by overlaying the circle of fifths (a circle representing
-pleasing intervals called fifths in music theory) onto the circle
-representing hue in HSV.  We establish the octave of the produced note
-using the value of the color in HSV, and the volume of the note using
-the saturation.
+In our transformation to music, we map colors (natively in RGB) to HSV
+and then to notes in MIDI.  We do this by overlaying the circle of
+fifths (a circle representing pleasing intervals called fifths in
+music theory) onto the circle representing hue in HSV.  We establish
+the octave of the produced note using the value of the color in HSV,
+and the volume of the note using the saturation.
 

### 2013-05-10 writeup/paper.tex
introducing HSV more clearly

@@
 data.  To do the conversion to tone, velocity pairs, we transform
-colors in the input data, meaning that we have to start with a
-representation of $R$ colors for each image.  To do this
+regions in the input data, meaning that we have to start with a
+representation of $n$ regions for each image.  To do this
 transformation, we considered using a MapReduce job: mapping over the

### 2013-05-10 writeup/paper.tex
fixups on explanation in paper for mapper

@@
 After taking the image data to an intermediate format, we must map the
-image data to a set of note / region pairs for use in the MIDI.  To do
-this, we map a conversion function across each of the different
-portions of each image.  The conversion happens by mapping over our
-set of images in the \code{Image2MusicMR} class.  The input of this
-reducer is a set of key/value pairs which constitute the image number
-and image data (a set of integers representing each pixel in the
-image).  The output of the image to music conversion is $n$ sets of
-note lists (where notes are expressed as pairs).  This is stored as a
-set of key/value pairs where the key is the region number and image
-number pair (subsequently discussed, but used for secondary sorting)
-and the part number and MIDI note.
+image data to a set of note / region pairs for use in producing MIDI
+output.  To do this, we map a conversion function across each region
+of each image.  The conversion happens by mapping over our set of
+images in the \code{Image2MusicMR} class.  The input of this mapper is
+a set of key/value pairs which constitute the image number and image
+data (a set of integers representing each pixel in the image).  The
+output of the image to music conversion is $n$ sets of note lists
+(where notes are expressed as pairs).  The reducer takes each part of
+the sore and combines it into a resulting MIDI file.  This is stored
+as a set of key/value pairs where the key is the region number and
+image number pair (subsequently discussed, but used for secondary
+sorting) and the part number and MIDI note.
 
@@
 collate together.  After mapping over each pixel, we need to end up
-with a stream of images for each region: a na\"{i}ve implementation simply
-maps each pixel in our input corpus to these tone, velocity pairs:
+with a stream of images for each region: a na\"{i}ve implementation
+simply maps each pixel in our input corpus to these tone, velocity
+pairs:
 

### 2013-05-10 writeup/paper.tex
small fixups

@@
 One might describe our final piece, \emph{Digital Nightmare}, as an unrelenting flood of almost patternless musical information.
-This characteristic emerges from some simplifications in design choices that a later vernon of our process would look to address.
+This characteristic emerges from some simplifications in design choices that a later version of our process would look to address.
 Here we present a list of such areas and possible solutions.
@@
   (beat). This results in constant note information without any space
-  for the listener to "catch up".  Darker colors below some threshold
-  of brightness could all be translated into musical rests.
+  for the listener to ``catch up.''  Darker colors below some
+  threshold of brightness could all be translated into musical rests.
 
@@
   only length note used.  Consecutive image regions with similar
-  colors could be "strung together" cause some instruments to play
+  colors could be ``strung together'' cause some instruments to play
   longer notes.
@@
 \item Late on in the process, we moved from saturation mapping to
-  velocity to choosing a fixed velocity of 127.  We found that
+  velocity to choosing a fixed velocity of $127$.  We found that
   velocity dynamics were not so perceptible with the degree of musical
